### 2025-03-30 docs/paper/figures/outputs-ascii-colored.tex
paper wip

\begin{Verbatim}[commandchars=\\\{\}]
# # # # # # # # # # #
#         \textcolor{blue}{X X X} #   #
#   # # # \textcolor{blue}{X} # \textcolor{blue}{X} #   #
#       # \textcolor{blue}{X} # \textcolor{green}{S}     #
# # # # # \textcolor{blue}{X} # # # # #
# \textcolor{blue}{X X X X X} # \textcolor{red}{E} \textcolor{blue}{X X} #
# \textcolor{blue}{X} # # #   # # # \textcolor{blue}{X} #
# \textcolor{blue}{X} #           # \textcolor{blue}{X} #
# \textcolor{blue}{X} # # # # # # # \textcolor{blue}{X} #
# \textcolor{blue}{X X X X X X X X X} #
# # # # # # # # # # #
\end{Verbatim}