### 2025-04-26 docs/paper/figures/tex/fig1_diagram.tex
move figure code to external files

\begin{figure} 
	\begin{minipage}{5in}
	  \input{diagram/diagram.tikz} 
	\end{minipage}
	\caption{
	  Usage of maze-dataset. We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}. This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats. Code in the image contains clickable links to \docslink{maze_dataset.html}{documentation}. A variety of generated examples can be viewed \docslink{examples/maze_examples.html}{here}.
	}
	\label{fig:diagram}
  \end{figure}