### 2025-05-28 docs/paper/preamble.tex

\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{calc}
\tikzset{ % Define a TikZ style for an external hyperlink node.
  hyperlink node url/.style={
	alias=sourcenode,
	append after command={
	  let \p1 = (sourcenode.north west),
		  \p2 = (sourcenode.south east),
		  \n1 = {\x2-\x1},
		  \n2 = {\y1-\y2} in
	  node[inner sep=0pt, outer sep=0pt, anchor=north west, at=(\p1)]
		  {\href{#1}{\XeTeXLinkBox{\phantom{\rule{\n1}{\n2}}}}}
	}
  }
}
\providecommand{\XeTeXLinkBox}[1]{#1}
\newcommand{\docslink}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{#2}}
\newcommand{\docslinkcode}[2]{\href{https://understanding-search.github.io/maze-dataset/#1}{\texttt{#2}}}
\newcommand{\docslinkcodemain}[1]{\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\##1}{\texttt{#1}}}
\newcommand{\secref}[1]{\hyperref[#1]{section: \textit{\nameref{#1}}}}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>0.8\textheight 0.8\textheight\else\Gin@nat@height\fi}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\makeatother