### 2025-05-28 docs/paper/figures/outputs-ascii-colored.tex
wip

{\ttfamily
\setlength{\tabcolsep}{0.5em}
\setlength{\arraystretch}{0.8}
\begin{tabular}{ccccccccccc}
\# & \# & \# & \# & \# & \# & \# & \# & \# & \# & \# \\
\# &   &   &   &   & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# &  & \# \\
\# &   & \# & \# & \# & \textcolor{blue}{X} & \# & \textcolor{blue}{X} & \# &   & \# \\
\# &   &   &   & \# & \textcolor{blue}{X} & \# & \textcolor{green}{S} &   &   & \# \\
\# & \# & \# & \# & \# & \textcolor{blue}{X} & \# & \# & \# & \# & \# \\
\# & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# & \textcolor{red}{E} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# \\
\# & \textcolor{blue}{X} & \# & \# & \# &   & \# & \# & \# & \textcolor{blue}{X} & \# \\
\# & \textcolor{blue}{X} & \# &   &   &   &   &   & \# & \textcolor{blue}{X} & \# \\
\# & \textcolor{blue}{X} & \# & \# & \# & \# & \# & \# & \# & \textcolor{blue}{X} & \# \\
\# & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \textcolor{blue}{X} & \# \\
\# & \# & \# & \# & \# & \# & \# & \# & \# & \# & \#
\end{tabular}
}