### 2025-07-28 docs/paper/figures/tex/fig4_tokenfmt.tex

\begin{figure} 
	\centering
	\hspace{-1em}
	\begin{minipage}{5in}
	  \footnotesize
	  \input{figures/outputs-tokens-colored.tex}
	\end{minipage}
	\caption{
	  Example text output format with token regions highlighted.
	  \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
	  \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
	  \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
	  \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence.
	  By passing an instance of \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#MazeTokenizerModular}{\texttt{MazeTokenizerModular}} to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#MazeDataset.as_tokens}{\texttt{as\_tokens(...)}} a maze can be converted to a text sequence. The \href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html#MazeTokenizerModular}{\texttt{MazeTokenizerModular}} class contains a rich set of options with 19 discrete parameters, resulting in over 5.8 million unique possible tokenizers.
	}
	\label{fig:token-regions}
\end{figure}