### 2025-08-04 docs/paper/figures/tex/fig6_benchmarks.tex
figure 6 fixed?

\begin{figure}
	\hypertarget{fig:benchmarks}{%
		\centering
		\includegraphics[width=0.9\textwidth,height=\textheight]{figures/benchmarks/gridsize-vs-gentime.pdf}
		\caption{
			Plot of maze generation time. Generation time scales
			exponentially with maze size for all algorithms. Generation time per
			maze does not depend on the number of mazes being generated, and there
			is minimal overhead to initializing the generation process for a small
			dataset. Wilson's algorithm is notably less efficient than others and
			has high variance. Note that values are averaged across all parameter
			sets for that algorithm. More information can be found on the
			\href{https://understanding-search.github.io/maze-dataset/benchmarks/}{benchmarks page}.
		}
		\label{fig:benchmarks}
	}
\end{figure}