### 2025-08-21 docs/paper/figures/tex/fig1_diagram.tex
fix minor typos

\begin{figure} 
	\resizebox{\linewidth}{!}{
    	\input{diagram/diagram.tikz}
	}
	\caption{
	  Usage of maze-dataset. We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}. This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats. A variety of generated examples can be viewed on the \docslink{examples/maze_examples.html}{examples page}, and more information can be found in the \docslink{maze_dataset.html}{documentation}.
	}
	\label{fig:diagram}
\end{figure}