### 2025-10-20 docs/paper/paper.tex
recompile paper

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage[rgb,svgnames]{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage[pdfa]{hyperref}
\usepackage{hyperxmp}
\hypersetup{%
    unicode=true,
    pdfapart=3,
    pdfaconformance=B,
    pdftitle={maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility},
    pdfauthor={Michael Igorevich Ivanitskiy, Aaron Sandoval, Alexander
F. Spies, Tilman Räuker, Brandon Knutson, Cecilia Diniz Behn, Samy Wu
Fung},
    pdfpublication={Journal of Open Source Software},
    pdfpublisher={Open Journals},
    pdfissn={2475-9066},
    pdfpubtype={journal},
    pdfvolumenum={},
    pdfissuenum={},
    pdfdoi={10.xxxxxx/draft},
    pdfcopyright={Copyright (c) 1970, Michael Igorevich Ivanitskiy,
Aaron Sandoval, Alexander F. Spies, Tilman Räuker, Brandon Knutson,
Cecilia Diniz Behn, Samy Wu Fung},
    pdflicenseurl={http://creativecommons.org/licenses/by/4.0/},
    colorlinks=true,
    linkcolor=[rgb]{0.0, 0.5, 1.0},
    citecolor=Blue,
    urlcolor=[rgb]{0.0, 0.5, 1.0},
    breaklinks=true
}
% https://tex.stackexchange.com/a/535849
% Create an OutputIntent in order to correctly specify colours
\immediate\pdfobj stream attr{/N 3} file{sRGB.icc}
\pdfcatalog{%
  /OutputIntents [
    <<
      /Type /OutputIntent
      /S /GTS_PDFA1
      /DestOutputProfile \the\pdflastobj\space 0 R
      /OutputConditionIdentifier (sRGB)
      /Info (sRGB)
    >>
  ]
}
\pdfvariable omitcidset=1
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{joss/logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Ivanitskiy
et al. (2025). maze-dataset: Maze Generation with Algorithmic Variety
and Representational Flexibility. \emph{Journal of Open Source
Software}, \emph{¿VOL?}(¿ISSUE?), ¿PAGE?
\url{https://doi.org/10.xxxxxx/draft}}.}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

\definecolor{c53baa1}{RGB}{83,186,161}
\definecolor{c202826}{RGB}{32,40,38}
\def \rorglobalscale {0.1}
\newcommand{\rorlogo}{%
\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364).. controls (3.9643, 1.4576) and (4.0508, 1.5081) .. (4.1098, 1.5878).. controls (4.169, 1.6674) and (4.1984, 1.7642) .. (4.1984, 1.8777).. controls (4.1984, 1.9719) and (4.182, 2.0503) .. (4.1495, 2.1132).. controls (4.1169, 2.1762) and (4.0727, 2.2262) .. (4.0174, 2.2635).. controls (3.9621, 2.3006) and (3.8976, 2.3273) .. (3.824, 2.3432).. controls (3.7505, 2.359) and (3.6727, 2.367) .. (3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688).. controls (2.9625, 1.8833) and (2.9572, 1.8976) .. (2.9514, 1.9119).. controls (2.9083, 2.0164) and (2.848, 2.1056) .. (2.7705, 2.1791).. controls (2.6929, 2.2527) and (2.6014, 2.3093) .. (2.495, 2.3487).. controls (2.3889, 2.3881) and (2.2728, 2.408) .. (2.1468, 2.408).. controls (2.0209, 2.408) and (1.905, 2.3881) .. (1.7986, 2.3487).. controls (1.6925, 2.3093) and (1.6007, 2.2527) .. (1.5232, 2.1791).. controls (1.4539, 2.1132) and (1.3983, 2.0346) .. (1.3565, 1.9436).. controls (1.3504, 2.009) and (1.3351, 2.0656) .. (1.3105, 2.1132).. controls (1.2779, 2.1762) and (1.2338, 2.2262) .. (1.1785, 2.2635).. controls (1.1232, 2.3006) and (1.0586, 2.3273) .. (0.985, 2.3432).. controls (0.9115, 2.359) and (0.8337, 2.367) .. (0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364).. controls (1.1255, 1.4576) and (1.2121, 1.5081) .. (1.2711, 1.5878).. controls (1.2737, 1.5915) and (1.2761, 1.5954) .. (1.2787, 1.5991).. controls (1.2782, 1.5867) and (1.2779, 1.5743) .. (1.2779, 1.5616).. controls (1.2779, 1.4327) and (1.2996, 1.3158) .. (1.3428, 1.2113).. controls (1.3859, 1.1068) and (1.4462, 1.0176) .. (1.5237, 0.944).. controls (1.601, 0.8705) and (1.6928, 0.8139) .. (1.7992, 0.7744).. controls (1.9053, 0.735) and (2.0214, 0.7152) .. (2.1474, 0.7152).. controls (2.2733, 0.7152) and (2.3892, 0.735) .. (2.4956, 0.7744).. controls (2.6016, 0.8139) and (2.6935, 0.8705) .. (2.771, 0.944).. controls (2.8482, 1.0176) and (2.9086, 1.1068) .. (2.952, 1.2113).. controls (2.9578, 1.2253) and (2.9631, 1.2398) .. (2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777).. controls (0.9438, 1.7534) and (0.92, 1.7357) .. (0.8911, 1.7243).. controls (0.8623, 1.7129) and (0.83, 1.706) .. (0.7945, 1.7039).. controls (0.7588, 1.7015) and (0.7252, 1.7005) .. (0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667).. controls (0.7477, 2.0667) and (0.7805, 2.0643) .. (0.8139, 2.0598).. controls (0.8472, 2.0553) and (0.8768, 2.0466) .. (0.9025, 2.0336).. controls (0.9282, 2.0206) and (0.9496, 2.0021) .. (0.9663, 1.9778).. controls (0.9829, 1.9534) and (0.9914, 1.9209) .. (0.9914, 1.8799).. controls (0.9914, 1.8362) and (0.9819, 1.8021) .. (0.9628, 1.7777) -- cycle(2.6125, 1.3533).. controls (2.5889, 1.2904) and (2.5553, 1.2359) .. (2.5112, 1.1896).. controls (2.4672, 1.1433) and (2.4146, 1.1073) .. (2.3529, 1.0814).. controls (2.2916, 1.0554) and (2.2228, 1.0427) .. (2.1471, 1.0427).. controls (2.0712, 1.0427) and (2.0026, 1.0557) .. (1.9412, 1.0814).. controls (1.8799, 1.107) and (1.8272, 1.1433) .. (1.783, 1.1896).. controls (1.7391, 1.2359) and (1.7052, 1.2904) .. (1.6817, 1.3533).. controls (1.6581, 1.4163) and (1.6465, 1.4856) .. (1.6465, 1.5616).. controls (1.6465, 1.6359) and (1.6581, 1.705) .. (1.6817, 1.7687).. controls (1.7052, 1.8325) and (1.7388, 1.8873) .. (1.783, 1.9336).. controls (1.8269, 1.9799) and (1.8796, 2.0159) .. (1.9412, 2.0418).. controls (2.0026, 2.0675) and (2.0712, 2.0804) .. (2.1471, 2.0804).. controls (2.223, 2.0804) and (2.2916, 2.0675) .. (2.3529, 2.0418).. controls (2.4143, 2.0161) and (2.467, 1.9799) .. (2.5112, 1.9336).. controls (2.5551, 1.8873) and (2.5889, 1.8322) .. (2.6125, 1.7687).. controls (2.636, 1.705) and (2.6477, 1.6359) .. (2.6477, 1.5616).. controls (2.6477, 1.4856) and (2.636, 1.4163) .. (2.6125, 1.3533) -- cycle(3.8015, 1.7777).. controls (3.7825, 1.7534) and (3.7587, 1.7357) .. (3.7298, 1.7243).. controls (3.701, 1.7129) and (3.6687, 1.706) .. (3.6333, 1.7039).. controls (3.5975, 1.7015) and (3.5639, 1.7005) .. (3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667).. controls (3.5864, 2.0667) and (3.6192, 2.0643) .. (3.6526, 2.0598).. controls (3.6859, 2.0553) and (3.7155, 2.0466) .. (3.7412, 2.0336).. controls (3.7669, 2.0206) and (3.7883, 2.0021) .. (3.805, 1.9778).. controls (3.8216, 1.9534) and (3.8301, 1.9209) .. (3.8301, 1.8799).. controls (3.8301, 1.8362) and (3.8206, 1.8021) .. (3.8015, 1.7777) -- cycle;
  \end{scope}
\end{tikzpicture}
}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\sffamily]{Ligatures=TeX}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

%% Font settings
\usepackage{fontsetup} % Lazy way to get proper Greek lowercase glyphs

% Use Hack https://sourcefoundry.org/hack/
\setmonofont{Hack}

\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\urlstyle{same}  % don't use monospace font for urls
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage{lineno}
\linenumbers
\usepackage{draftwatermark}

\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\input{preamble.tex}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{maze-dataset: Maze Generation with Algorithmic Variety and
Representational Flexibility}

\author[1%
%
\ensuremath\mathparagraph]{Michael Igorevich Ivanitskiy%
  \,\orcidlink{0000-0002-4213-4993}\,%
}
\author[4%
%
]{Aaron Sandoval%
  \,\orcidlink{0009-0002-8380-6140}\,%
}
\author[2%
%
]{Alexander F. Spies%
  \,\orcidlink{0000-0002-8708-1530}\,%
}
\author[3%
%
]{Tilman Räuker%
  \,\orcidlink{0009-0009-6321-4413}\,%
}
\author[1%
%
]{Brandon Knutson%
  \,\orcidlink{0009-0004-8413-0239}\,%
}
\author[1%
%
]{Cecilia Diniz Behn%
  \,\orcidlink{0000-0002-8078-5105}\,%
}
\author[1%
%
]{Samy Wu Fung%
  \,\orcidlink{0000-0002-2926-4582}\,%
}

\affil[1]{Colorado School of Mines, Department of Applied Mathematics
and Statistics%
}
\affil[2]{Imperial College London%
}
\affil[3]{UnSearch.org%
}
\affil[4]{Independent%
}
\affil[$\mathparagraph$]{Corresponding author}
\date{\vspace{-2.5ex}}

\begin{document}
\maketitle

\marginpar{

  \begin{flushleft}
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/10.xxxxxx/draft}{\color{linky}{10.xxxxxx/draft}}

  \vspace{2mm}
    {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{https://github.com/openjournals}{\color{linky}{Review}} \ExternalLink
    \item \href{https://github.com/openjournals}{\color{linky}{Repository}} \ExternalLink
    \item \href{https://doi.org/10.5281}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

    \par\noindent\hrulefill\par

  \vspace{2mm}

  {\bfseries Editor:} \href{https://joss.theoj.org}{Open
Journals} \ExternalLink
   \\
  \vspace{1mm}
    {\bfseries Reviewers:}
  \begin{itemize}
  \setlength\itemsep{0em}
    \item \href{https://github.com/openjournals}{@openjournals}
    \end{itemize}
    \vspace{2mm}

    {\bfseries Submitted:} 01 January 1970\\
    {\bfseries Published:} unpublished

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{https://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).

  \end{flushleft}
}

\section{Summary}\label{summary}

Solving mazes is a classic problem in computer science and artificial
intelligence, and humans have been constructing mazes for thousands of
years. Although finding the shortest path through a maze is a solved
problem, this very fact makes it an excellent testbed for studying how
machine learning algorithms solve problems and represent spatial
information. We introduce \texttt{maze-dataset}, a user-friendly Python
library for generating, processing, and visualizing datasets of mazes.
This library supports a variety of maze generation algorithms which can
be configured with various parameters, and the resulting mazes can be
filtered to satisfy desired properties. Also provided are tools for
converting mazes to and from various formats suitable for a variety of
neural network architectures, such as rasterized images, tokenized text
sequences, and various visualizations. As well as providing a simple
interface for generating, storing, and loading these datasets,
\texttt{maze-dataset} is extensively tested, type hinted, benchmarked,
and documented.

\input{figures/tex/fig1_diagram.tex}

\section{Statement of Need}\label{statement-of-need}

While maze generation itself is straightforward, the architectural
challenge comes from building a system supporting many algorithms with
configurable parameters, property filtering, representation
transformation, and reproducibility. This library aims to greatly
streamline the process of generating and working with datasets of mazes
that can be described as subgraphs of an \(n \times n\) lattice with
boolean connections and, optionally, start and end points that are nodes
in the graph. Furthermore, we place emphasis on a wide variety of
possible text output formats aimed at evaluating the spatial reasoning
capabilities of Large Language Models (LLMs) and other text-based
transformer models.

For interpretability and behavioral research, algorithmic tasks offer
benefits by allowing systematic data generation and task decomposition,
as well as simplifying the process of circuit discovery
(\citeproc{ref-interpretability-survey}{Räuker et al., 2023}). Although
mazes are well suited for these investigations, we found that existing
maze generation packages (\citeproc{ref-cobbe2019procgen}{Cobbe et al.,
2019}; \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
\citeproc{ref-harriesMazeExplorerCustomisable3D2019}{Harries et al.,
n.d.}; \citeproc{ref-gh_Nemeth_2019}{Németh, 2019};
\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal, et
al., 2021}) lack support for transforming between multiple
representations and provide limited control over the maze generation
process.

\subsection{Related Works}\label{related-works}

A multitude of public and open-source software packages exist for
generating mazes (\citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
\citeproc{ref-gh_Nemeth_2019}{Németh, 2019};
\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal, et
al., 2021}). However, nearly all of these packages produce mazes
represented as rasterized images or other visual formats rather than the
underlying graph structure, and this makes it difficult to work with
these datasets.

\begin{itemize}
\item
  Most prior works provide mazes in visual or raster formats, and we
  provide a variety of similar output formats:

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}
    (\citeproc{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta, Bansal,
    et al., 2021}) and can be configured to be similar to the outputs of
    Németh (\citeproc{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to Singla
    (\citeproc{ref-eval-gpt-visual}{2023}) and Oppenheim
    (\citeproc{ref-gh-oppenheimj2018maze}{2018})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of many packages (\citeproc{ref-mazegenerator-net}{Alance
    AB, 2019}; \citeproc{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \citeproc{ref-mathematica-maze}{Guo et al., 2011};
    \citeproc{ref-mdl-suite}{Nag, 2020}).
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of Liu \& Wu (\citeproc{ref-eval-LLM-graphs}{2023})
  but with many more options, detailed in
  \hyperref[sec:tokenized-output-formats]{section: \textit{\nameref{sec:tokenized-output-formats}}}.
\item
  Preserving metadata about the generation algorithm with the dataset
  itself is essential for studying the effects of distributional shifts.
  Our package efficiently stores the dataset along with its metadata in
  a single human-readable file (\citeproc{ref-zanj}{M. Ivanitskiy,
  n.d.}). As far as we are aware, no existing packages do this reliably.
\item
  Storing mazes as images or adjacency matrices is not only difficult to
  work with, but also inefficient. We use a highly efficient method
  detailed in
  \hyperref[sec:implementation]{section: \textit{\nameref{sec:implementation}}}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\citeproc{ref-ayaz2008maze}{Ayaz et al., 2008})
  appear to simply no longer be accessible.
\end{itemize}

\newpage

\section{Features}\label{features}

We direct readers to our
\href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{examples},
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{docs},
and
\href{https://understanding-search.github.io/maze-dataset/notebooks/}{notebooks}
for more information. Our package can be installed from
\href{https://pypi.org/project/maze-dataset/}{PyPi} via
\texttt{pip\ install\ maze-dataset}, or directly from the
\href{https://github.com/understanding-search/maze-dataset}{git
repository} (\citeproc{ref-maze-dataset-github}{Michael I. Ivanitskiy et
al., 2023a}).

Datasets of mazes are created from a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
configuration object, which allows specifying the number of mazes, their
size, the generation algorithm, and various parameters for the
generation algorithm. Datasets can also be filtered after generation to
satisfy certain properties. Custom filters can be specified, and some
filters are included in
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/filters.html\#MazeDatasetFilters}{\texttt{MazeDatasetFilters}}.

\subsection{Visual Output Formats}\label{visual-output-formats}

Internally, mazes are
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#SolvedMaze}{\texttt{SolvedMaze}}
objects, which have path information and an array optimized for storing
sub-graphs of a lattice. These objects can be converted to and from
several formats, shown in \autoref{fig:output-fmts}, to maximize their
utility in different contexts.

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\citeproc{ref-deepthinking}{Schwarzschild, Borgnia, Gupta, Huang, et
al., 2021}). To facilitate the use of our package in this context, we
replicate the format of Schwarzschild, Borgnia, Gupta, Bansal, et al.
(\citeproc{ref-easy_to_hard}{2021}) and provide the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
class which returns rasterized pairs of (input, target) mazes as shown
in \autoref{fig:e2h-raster}.

\input{figures/tex/fig2_formats.tex}

\input{figures/tex/fig3_raster.tex}

\newpage

\subsection{Tokenized Output
Formats}\label{sec:tokenized-output-formats}

Autoregressive transformer models can be quite sensitive to the exact
format of input data, and may even use delimiter tokens to perform
reasoning steps (\citeproc{ref-pfau2024dotbydot}{Pfau et al., 2024};
\citeproc{ref-spies2024causalworldmodels}{Spies et al., 2024}). To
facilitate systematic investigation of the effects of different
representations of data on text model performance, we provide a variety
of text output formats, with an example given in
\autoref{fig:token-regions}. We utilize Finite State Transducers
(\citeproc{ref-Gallant2015Transducers}{Gallant, 2015}) for efficiently
storing valid tokenizers.

\input{figures/tex/fig4_tokenfmt.tex}

\section{Implementation}\label{sec:implementation}

Using an adjacency matrix for storing mazes would be memory inefficient
by failing to exploit the highly sparse structure, while using an
adjacency list could lead to a poor lookup time. This package utilizes a
simple, efficient representation of mazes as subgraphs of a finite
lattice, detailed in \autoref{fig:maze-impl}, which we call a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.

\input{figures/tex/fig8_impl.tex}

Our package is implemented in Python (\citeproc{ref-python}{Rossum,
1995}), and makes use of the extensive scientific computing ecosystem,
including NumPy (\citeproc{ref-numpy}{Harris et al., 2020}) for array
manipulation, plotting tools (\citeproc{ref-matplotlib}{Hunter, 2007};
\citeproc{ref-seaborn}{Waskom, 2021}), Jupyter notebooks
(\citeproc{ref-jupyter}{Kluyver et al., 2016}), and PySR
(\citeproc{ref-pysr}{Cranmer, 2023}) for symbolic regression.

\subsection{Benchmarks}\label{benchmarks}

We benchmarks for generation time across various configurations in
\autoref{tab:benchmarks} and \autoref{fig:benchmarks}. Experiments were
performed on a
\href{https://docs.github.com/en/actions/using-github-hosted-runners/using-github-hosted-runners/about-github-hosted-runners\#standard-github-hosted-runners-for-public-repositories}{standard
GitHub runner} without parallelism. Additionally, maze generation under
certain constraints may not always be successful, and for this we
provide a way to estimate the success rate of a given configuration,
described in \autoref{fig:sre}.

\input{figures/tex/tab1_benchmarks.tex}

\input{figures/tex/fig6_benchmarks.tex}

\input{figures/tex/fig7_sre.tex}

\section{Usage in Research}\label{usage-in-research}

This package was originally built for the needs of the
\texttt{maze-transformer} project
(\citeproc{ref-maze-transformer-github}{Michael I. Ivanitskiy et al.,
2023b}), which aims to investigate spatial planning and world models in
autoregressive transformer models trained on mazes
(\citeproc{ref-ivanitskiy2023structuredworldreps}{Michael Igorevich
Ivanitskiy, Spies, et al., 2023};
\citeproc{ref-maze-dataset-arxiv-2023}{Michael Igorevich Ivanitskiy,
Shah, et al., 2023}; \citeproc{ref-spies2024causalworldmodels}{Spies et
al., 2024}). It was extended for work on understanding the mechanisms by
which recurrent convolutional and implicit networks
(\citeproc{ref-fung2022jfb}{Fung et al., 2022}) solve mazes given a
rasterized view (\citeproc{ref-knutson2024logicalextrapolation}{Knutson
et al., 2024}), which required matching the pixel-padded and endpoint
constrained output format of Schwarzschild, Borgnia, Gupta, Bansal, et
al. (\citeproc{ref-easy_to_hard}{2021}). Ongoing work using
\texttt{maze-dataset} aims to investigate the effects of varying the
tokenization format on the performance of pretrained LLMs on spatial
reasoning.

At the time of writing, this software package has been actively used in
work by other groups:

\begin{itemize}
\item
  By Nolte et al. (\citeproc{ref-nolte2024multistep}{2024}) to compare
  the effectiveness of transformers trained with the MLM-\(\mathcal{U}\)
  (\citeproc{ref-MLMU-kitouni2024factorization}{Kitouni et al., 2024})
  multistep prediction objective against standard autoregressive
  training for multi-step planning on our maze task.
\item
  By Wang et al. (\citeproc{ref-wang2024imperative}{2024}) and Chen et
  al. (\citeproc{ref-chen2024iaimperative}{2024}) to study imperative
  learning.
\item
  By Zhang et al. (\citeproc{ref-zhang2025tscend}{2025a}) to introduce a
  novel framework for reasoning diffusion models.
\item
  By Dao \& Vu (\citeproc{ref-dao2025alphamaze}{2025}) to improve
  spatial reasoning in LLMs with GRPO.
\item
  By Cai et al. (\citeproc{ref-cai2025morse}{2025}) to create a
  multimodal reasoning benchmark, via mazes in videos.
\item
  By Xu et al. (\citeproc{ref-xu2025visual}{2025}) to study visual
  planning in LLMs.
\item
  By Lee et al. (\citeproc{ref-lee2025adaptive}{2025}) to evaluate
  adaptive inference-time scaling with diffusion models on maze
  navigation tasks.
\item
  By Zhang et al. (\citeproc{ref-zhang2025vfscale}{2025b}) to test
  verifier-free diffusion models.
\end{itemize}

\newpage

\section{Acknowledgements}\label{acknowledgements}

\input{figures/tex/acknowledgements.tex}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0.5}
\bibitem[\citeproctext]{ref-mazegenerator-net}
Alance AB. (2019). \emph{Maze generator}.
\url{http://www.mazegenerator.net}.

\bibitem[\citeproctext]{ref-ayaz2008maze}
Ayaz, H., Allen, S. L., Platek, S. M., \& Onaral, B. (2008). Maze suite
1.0: A complete set of tools to prepare, present, and analyze
navigational and spatial cognitive neuroscience experiments.
\emph{Behavior Research Methods}, \emph{40}, 353--359.
\url{https://doi.org/10.3758/brm.40.1.353}

\bibitem[\citeproctext]{ref-cai2025morse}
Cai, Z., Wang, A., Satheesh, A., Nakhawa, A., Jae, H., Powell, K., Liu,
M., Jay, N., Oh, S., Wang, X., \& others. (2025). MORSE-500: A
programmatically controllable video benchmark to stress-test multimodal
reasoning. \emph{arXiv Preprint arXiv:2506.05523}.
\url{https://doi.org/10.48550/arXiv.2506.05523}

\bibitem[\citeproctext]{ref-chen2024iaimperative}
Chen, X., Yang, F., \& Wang, C. (2024). {iA*}: Imperative learning-based
{A*} search for pathfinding. \emph{arXiv Preprint arXiv:2403.15870}.
\url{https://doi.org/10.48550/arXiv.2403.15870}

\bibitem[\citeproctext]{ref-cobbe2019procgen}
Cobbe, K., Hesse, C., Hilton, J., \& Schulman, J. (2019). Leveraging
procedural generation to benchmark reinforcement learning. \emph{arXiv
Preprint arXiv:1912.01588}.
\url{https://doi.org/10.48550/arXiv.1912.01588}

\bibitem[\citeproctext]{ref-pysr}
Cranmer, M. (2023). Interpretable machine learning for science with PySR
and SymbolicRegression.jl. \emph{arXiv Preprint arXiv:2305.01582}.
\url{https://doi.org/10.48550/arXiv.2305.01582}

\bibitem[\citeproctext]{ref-dao2025alphamaze}
Dao, A., \& Vu, D. B. (2025). AlphaMaze: Enhancing large language
models' spatial intelligence via GRPO. \emph{arXiv Preprint
arXiv:2502.14669}. \url{https://doi.org/10.48550/arXiv.2502.14669}

\bibitem[\citeproctext]{ref-gh_Ehsan_2022}
Ehsan, E. (2022). \emph{Maze}. \url{https://github.com/emadehsan/maze}

\bibitem[\citeproctext]{ref-fung2022jfb}
Fung, S. W., Heaton, H., Li, Q., McKenzie, D., Osher, S., \& Yin, W.
(2022). Jfb: Jacobian-free backpropagation for implicit networks.
\emph{Proceedings of the AAAI Conference on Artificial Intelligence},
\emph{36}, 6648--6656. \url{https://doi.org/10.1609/aaai.v36i6.20619}

\bibitem[\citeproctext]{ref-Gallant2015Transducers}
Gallant, A. (2015). \emph{Index 1,600,000,000 keys with automata and
rust}. \url{https://burntsushi.net/transducers/}.

\bibitem[\citeproctext]{ref-mathematica-maze}
Guo, C., Barthelet, L., \& Morris, R. (2011). \emph{Maze generator and
solver}. Wolfram Demonstrations Project,
\url{https://demonstrations.wolfram.com/MazeGeneratorAndSolver/}.

\bibitem[\citeproctext]{ref-harriesMazeExplorerCustomisable3D2019}
Harries, L., Lee, S., Rzepecki, J., Hofmann, K., \& Devlin, S. (n.d.).
{MazeExplorer}: {A Customisable 3D Benchmark} for {Assessing
Generalisation} in {Reinforcement Learning}. \emph{2019 {IEEE Conf}.
{Games CoG}}, 1--4. \url{https://doi.org/10.1109/cig.2019.8848048}

\bibitem[\citeproctext]{ref-numpy}
Harris, C. R., Millman, K. J., Walt, S. J. van der, Gommers, R.,
Virtanen, P., \& al., et. (2020). Array programming with NumPy.
\emph{Nature}, \emph{585}, 357--362.
\url{https://doi.org/10.1038/s41586-020-2649-2}

\bibitem[\citeproctext]{ref-matplotlib}
Hunter, J. D. (2007). Matplotlib: A 2D graphics environment.
\emph{Computing in Science and Engineering}, \emph{9}(3), 90--95.
\url{https://doi.org/10.1109/MCSE.2007.55}

\bibitem[\citeproctext]{ref-zanj}
Ivanitskiy, M. (n.d.). \emph{ZANJ}.
\url{https://doi.org/10.5281/zenodo.15540393}

\bibitem[\citeproctext]{ref-maze-dataset-github}
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023a).
\emph{Maze dataset}. \url{https://doi.org/10.48550/arXiv.2309.10498}

\bibitem[\citeproctext]{ref-maze-transformer-github}
Ivanitskiy, Michael I., Shah, R., Spies, A. F., Räuker, T., Valentine,
D., Rager, C., Quirke, L., Corlouer, G., \& Mathwin, C. (2023b).
\emph{Maze transformer interpretability}.
\url{https://doi.org/10.48550/arXiv.2312.02566}

\bibitem[\citeproctext]{ref-maze-dataset-arxiv-2023}
Ivanitskiy, Michael Igorevich, Shah, R., Spies, A. F., Räuker, T.,
Valentine, D., Rager, C., Quirke, L., Mathwin, C., Corlouer, G., Behn,
C. D., \& others. (2023). A configurable library for generating and
manipulating maze datasets. \emph{arXiv Preprint arXiv:2309.10498}.
\url{https://doi.org/10.48550/arXiv.2309.10498}

\bibitem[\citeproctext]{ref-ivanitskiy2023structuredworldreps}
Ivanitskiy, Michael Igorevich, Spies, A. F., Räuker, T., Corlouer, G.,
Mathwin, C., Quirke, L., Rager, C., Shah, R., Valentine, D., Behn, C.
D., \& others. (2023). Structured world representations in maze-solving
transformers. \emph{arXiv Preprint arXiv:2312.02566}.
\url{https://doi.org/10.48550/arXiv.2312.02566}

\bibitem[\citeproctext]{ref-MLMU-kitouni2024factorization}
Kitouni, O., Nolte, N. S., Williams, A., Rabbat, M., Bouchacourt, D., \&
Ibrahim, M. (2024). The factorization curse: Which tokens you predict
underlie the reversal curse and more. \emph{Advances in Neural
Information Processing Systems}, \emph{37}, 112329--112355.
\url{https://doi.org/10.48550/arXiv.2406.05183}

\bibitem[\citeproctext]{ref-jupyter}
Kluyver, T., Ragan-Kelley, B., Perez, F., Granger, B., Bussonnier, M.,
Frederic, J., Kelley, K., Hamrick, J., Grout, J., Corlay, S., Ivanov,
P., Avila, D., Abdalla, S., \& Willing, C. (2016). Jupyter notebooks - a
publishing format for reproducible computational workflows.
\emph{Proceedings of the 20th International Conference on Electronic
Publishing}, 87--90. \url{https://doi.org/10.3233/978-1-61499-649-1-87}

\bibitem[\citeproctext]{ref-knutson2024logicalextrapolation}
Knutson, B., Rabeendran, A. C., Ivanitskiy, M., Pettyjohn, J.,
Diniz-Behn, C., Fung, S. W., \& McKenzie, D. (2024). On logical
extrapolation for mazes with recurrent and implicit networks.
\emph{arXiv Preprint arXiv:2410.03020}.
\url{https://doi.org/10.48550/arXiv.2410.03020}

\bibitem[\citeproctext]{ref-lee2025adaptive}
Lee, G., Bao, T. N. N., Yoon, J., Lee, D., Kim, M., Bengio, Y., \& Ahn,
S. (2025). Adaptive cyclic diffusion for inference scaling. \emph{arXiv
Preprint arXiv:2505.14036}.
\url{https://doi.org/10.48550/arXiv.2505.14036}

\bibitem[\citeproctext]{ref-eval-LLM-graphs}
Liu, C., \& Wu, B. (2023). Evaluating large language models on graphs:
Performance insights and comparative analysis. \emph{arXiv Preprint
arXiv:2308.11224}. \url{https://doi.org/10.48550/arXiv.2308.11224}

\bibitem[\citeproctext]{ref-mdl-suite}
Nag, A. (2020). MDL suite: A language, generator and compiler for
describing mazes. \emph{Journal of Open Source Software}, \emph{5}(46),
1815. \url{https://doi.org/10.21105/joss.01815}

\bibitem[\citeproctext]{ref-gh_Nemeth_2019}
Németh, F. (2019). \emph{Maze-generation-algorithms}.
\url{https://github.com/ferenc-nemeth/maze-generation-algorithms}

\bibitem[\citeproctext]{ref-nolte2024multistep}
Nolte, N., Kitouni, O., Williams, A., Rabbat, M., \& Ibrahim, M. (2024).
Transformers can navigate mazes with multi-step prediction. \emph{arXiv
Preprint arXiv:2412.05117}.
\url{https://doi.org/10.48550/arXiv.2412.05117}

\bibitem[\citeproctext]{ref-gh-oppenheimj2018maze}
Oppenheim, J. (2018). \emph{Maze-generator: Generate a random maze
represented as a 2D array using depth-first search}.
\url{https://github.com/oppenheimj/maze-generator/}; GitHub.

\bibitem[\citeproctext]{ref-pfau2024dotbydot}
Pfau, J., Merrill, W., \& Bowman, S. R. (2024). Let's think dot by dot:
Hidden computation in transformer language models. \emph{arXiv Preprint
arXiv:2404.15758}. \url{https://doi.org/10.48550/arXiv.2404.15758}

\bibitem[\citeproctext]{ref-interpretability-survey}
Räuker, T., Ho, A., Casper, S., \& Hadfield-Menell, D. (2023). Toward
transparent ai: A survey on interpreting the inner structures of deep
neural networks. \emph{2023 IEEE Conference on Secure and Trustworthy
Machine Learning (SaTML)}, 464--483.
\url{https://doi.org/10.1109/satml54575.2023.00039}

\bibitem[\citeproctext]{ref-python}
Rossum, G. van. (1995). \emph{Python reference manual} (CS-R9525).
Centrum voor Wiskunde; Informatica (CWI).
\url{https://ir.cwi.nl/pub/5008/05008D.pdf}

\bibitem[\citeproctext]{ref-easy_to_hard}
Schwarzschild, A., Borgnia, E., Gupta, A., Bansal, A., Emam, Z., Huang,
F., Goldblum, M., \& Goldstein, T. (2021). \emph{Datasets for {Studying
Generalization} from {Easy} to {Hard Examples}} (No. arXiv:2108.06011).
{arXiv}. \url{https://doi.org/10.48550/arXiv.2108.06011}

\bibitem[\citeproctext]{ref-deepthinking}
Schwarzschild, A., Borgnia, E., Gupta, A., Huang, F., Vishkin, U.,
Goldblum, M., \& Goldstein, T. (2021). Can you learn an algorithm?
Generalizing from easy to hard problems with recurrent networks.
\emph{Advances in Neural Information Processing Systems}, \emph{34},
6695--6706. \url{https://doi.org/10.48550/arXiv.2106.04537}

\bibitem[\citeproctext]{ref-eval-gpt-visual}
Singla, A. (2023). Evaluating ChatGPT and GPT-4 for visual programming.
\emph{arXiv Preprint arXiv:2308.02522}.
\url{https://doi.org/10.48550/arXiv.2308.02522}

\bibitem[\citeproctext]{ref-spies2024causalworldmodels}
Spies, A. F., Edwards, W., Ivanitskiy, M. I., Skapars, A., Räuker, T.,
Inoue, K., Russo, A., \& Shanahan, M. (2024). Transformers use causal
world models in maze-solving tasks. \emph{arXiv Preprint
arXiv:2412.11867}. \url{https://doi.org/10.48550/arXiv.2412.11867}

\bibitem[\citeproctext]{ref-wang2024imperative}
Wang, C., Ji, K., Geng, J., Ren, Z., Fu, T., Yang, F., Guo, Y., He, H.,
Chen, X., Zhan, Z., \& others. (2024). Imperative learning: A
self-supervised neural-symbolic learning framework for robot autonomy.
\emph{arXiv Preprint arXiv:2406.16087}.
\url{https://doi.org/10.48550/arXiv.2406.16087}

\bibitem[\citeproctext]{ref-seaborn}
Waskom, M. L. (2021). {seaborn}: Statistical data visualization.
\emph{Journal of Open Source Software}, \emph{6}(60), 3021.
\url{https://doi.org/10.21105/joss.03021}

\bibitem[\citeproctext]{ref-xu2025visual}
Xu, Y., Li, C., Zhou, H., Wan, X., Zhang, C., Korhonen, A., \& Vulić, I.
(2025). Visual planning: Let's think only with images. \emph{arXiv
Preprint arXiv:2505.11409}.
\url{https://doi.org/10.48550/arXiv.2505.11409}

\bibitem[\citeproctext]{ref-zhang2025tscend}
Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025a). T-SCEND: Test-time
scalable MCTS-enhanced diffusion model. \emph{arXiv Preprint
arXiv:2502.01989}. \url{https://doi.org/10.48550/arXiv.2502.01989}

\bibitem[\citeproctext]{ref-zhang2025vfscale}
Zhang, T., Pan, J.-S., Feng, R., \& Wu, T. (2025b). VFScale: Intrinsic
reasoning through verifier-free test-time scalable diffusion model.
\emph{arXiv Preprint arXiv:2502.01989}.
\url{https://doi.org/10.48550/arXiv.2502.01989}

\end{CSLReferences}

\end{document}